---
title: "Models for urban acoustics"
author: "Fiona Spooner"
date: "4 September 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

Things that still need to be addressed

* Invert auc_an_min_bio so that positive values mean a positive message!
* Account for date/weather in the models some how. Have min/mean/max temp for the overall period but something more finescale could be useful, especially precipitation as this will affect noise. 
* Should integrate all the preceding code into this file
* Open/Garden data missing from some of the sites ~ 12 in the 100m versions

```{r, eval = TRUE, message=FALSE, warning = FALSE}
library(here)
library(dplyr)
library(reshape2)
library(ggplot2)
library(DescTools)
library(knitr)
library(dplyr)
library(lubridate)
library(pracma)
library(ordinal)
library(MASS)
library(ppcor)
library(lme4)
library(bbmle)

```

```{r}
xy<-read.csv(here::here("allSitesDetailsCoordinates.csv"))

xy$SiteTypeRgp <- as.character(xy$Site_Type)
xy$SiteTypeRgp[xy$Site_Type == "Allotment"] <- "Community Garden"
xy$SiteTypeRgp[xy$Site_Type == "Green Wall" | xy$Site_Type == "Green Roof"] <- "Green Infrastructure"


hab<-read.csv(here::here("allSitesHabitat.csv"))
area<-read.csv(here::here("siteAreaAndPerimeter.csv"))
temp<-read.csv(here::here("temperature.csv"))
garden<-read.csv(here::here("garden100m.csv"))
colnames(garden)[3]<-"Area_Garden"
open<-read.csv(here::here("openSpace100m.csv"))
colnames(open)[3]<-"Area_Open"

lidar <- read.csv(here::here("100m_results.csv"))
lidar_cell_area <- 8.105 * 8.105
radius <- 100

lidar$siteName <- gsub("_", "", lidar$siteName)
lidar$grassProp <- (lidar$grassArea * lidar_cell_area)/(pi * radius^2)
lidar$shrubProp <-  (lidar$shrubArea * lidar_cell_area)/(pi * radius^2)
lidar$treeProp <-  (lidar$treeArea * lidar_cell_area)/(pi * radius^2)

df<-readRDS("all_days_feature_extractions_2019_09_02.RDS")

df<-df %>% 
      dplyr::select(site, day, steepest_anth, steepest_bio, steepest_time_anth, steepest_time_bio, auc_anthro, auc_bio, n_peaks_anth, n_peaks_bio, auc_an_min_bio, peak_anth, peak_bio, peak_anth_time, peak_bio_time, steepest_dawn_anth, steepest_dawn_bio, auc_dawn_anthro, auc_dawn_bio, dawn_auc_an_min_bio, peak_dawn_anth, peak_dawn_bio, steepest_dusk_anth, steepest_dusk_bio, auc_dusk_anthro, auc_dusk_bio, dusk_auc_an_min_bio, peak_dusk_anth, peak_dusk_bio) %>% 
    distinct()

df_hab <- merge(df,hab, by.x = "site", by.y="Site" )
df_hab <- merge(df_hab,area ,by.x = "site", by.y = "SiteCode")
df_hab <- merge(df_hab, temp, by.x = "site", by.y = "Site")
df_hab <- merge(df_hab, xy[,c("SiteCode", "SiteTypeRgp","StartDate", "EndDate")], by.x = "site", by.y = "SiteCode")
df_hab <- merge(df_hab, open[,c(2,3)], by.x = "site", by.y = "SiteCode", all = TRUE)
df_hab <- merge(df_hab, garden[,c(2,3)], by.x = "site", by.y = "SiteCode", all = TRUE)
df_hab <- merge(df_hab, lidar, by.x = "site", by.y = "siteName", all = TRUE)

df_hab$n_peaks_anth[is.na(df_hab$n_peaks_anth)]<-0
df_hab$n_peaks_bio[is.na(df_hab$n_peaks_bio)]<-0

df_hab$tempRange<-df_hab$maxTemp - df_hab$minTemp



```


Remove start day and end day as these are incomplete days
```{r}

df_hab$StartDate<-as.Date(df_hab$StartDate, format = "%d-%m-%y")
df_hab$EndDate<-as.Date(df_hab$EndDate, format = "%d-%m-%y")

df_hab<-df_hab %>% 
          group_by(site) %>% 
          filter(day != day(StartDate) & day != day(EndDate)) %>% 
          ungroup()
```


```{r}

df_hab$month <- as.factor(month(df_hab$EndDate))
df_hab$total_grn <- df_hab$treeProp + df_hab$shrubProp + df_hab$grassProp


ggplot(df_hab, aes(x = total_grn, y = auc_bio, group = site,colour = month ))+
  geom_point(width = 0.2)
```

```{r}

ggplot(df_hab, aes(x = total_grn, y = auc_anthro, group = site,colour = month))+
  geom_point(width = 0.2)

```


```{r}

ggplot(df_hab, aes(x = total_grn , y = peak_bio, group = site,colour =  month))+
  geom_point()

```
```{r}

ggplot(df_hab, aes(x = total_grn , y = peak_anth, group = site,colour =  month))+
  geom_point()

```

```{r}

ggplot(df_hab, aes(x = grassArea, y = auc_bio, group = site,colour = site))+
  geom_jitter(width = 0.2)+ 
  theme(legend.position = "none")

```

```{r}

ggplot(df_hab, aes(x = grassArea, y = auc_anthro, group = site,colour = site))+
  geom_jitter(width = 0.2)+ 
  theme(legend.position = "none")

```


```{r}

ggplot(df_hab, aes(x = Complexity, y = auc_an_min_bio, group = site,colour = site))+
  geom_jitter(width = 0.2)+ 
  theme(legend.position = "none")

```

```{r}

ggplot(df_hab, aes(x = siteDiversity, y = auc_bio, colour = Site_Type))+
  geom_jitter(width = 0.2)+
  theme(legend.position = "none")

```
```{r}
ggplot(df_hab, aes(x = siteDiversity, y = auc_anthro, colour = Site_Type))+
  geom_jitter(width = 0.2)+
  theme(legend.position = "none")

```


```{r}

ggplot(df_hab, aes(x = Area, y = auc_bio, colour = Site_Type))+
  geom_jitter(width = 0.2)+
  theme(legend.position = "none")

```


```{r}
ggplot(df_hab, aes(x =Area_Open, y = auc_bio, colour = site))+
  geom_jitter(width = 0.2)+
  theme(legend.position = "none")

```

```{r}
ggplot(df_hab, aes(x =Area_Garden, y = auc_bio, colour = site))+
  geom_jitter(width = 0.2)+
  theme(legend.position = "none")

```

```{r}
ggplot(df_hab, aes(x =Area_Open+Area_Garden, y = auc_bio, col=  Site_Type))+
  #geom_smooth()+
  geom_jitter(width = 0.2)

```

```{r}
ggplot(df_hab, aes(x =Area_Open+Area_Garden, y = peak_bio))+
  geom_smooth()+
  ylim(0,1)+
  geom_jitter(width = 0.2)+
  theme(legend.position = "none")

```


```{r}
ggplot(df_hab, aes(x =Area_Open+Area_Garden, y = peak_anth))+
  geom_smooth()+
  ylim(0,1)+
  geom_jitter(width = 0.2)+
  theme(legend.position = "none")

```

```{r, eval = FALSE, echo = FALSE}


# df_hab<-df_hab[complete.cases(df_hab),]
# 
# pcor(df_hab[,c(3,4,7:13,16:29)],method = "pearson")
# 
# cors<-pcor(df_hab[,c(7:10)],method = "pearson")
# 
# cors$p.value < 0.05
# 
# pcor(df_hab[,c(7:13,16:29)],method = "pearson")
# 
# 

```


```{r}

plot(df_hab[,c(9:13,21,22,36,43)])

# vars<-colnames(df_hab[,c(9:13,21,22,36,43)])   #better to name these
# df_hab_sc<-scale(df_hab[,vars])


df_hab_sel <- scale(df_hab %>% 
  dplyr::select(auc_anthro:peak_bio, peak_dawn_anth, peak_dawn_bio))


df_hab_sc<-data.frame(Complexity = df_hab$Complexity, siteDiversity = df_hab$siteDiversity, site = df_hab$site,day = df_hab$day,Area_Garden = df_hab$Area_Garden, Area_Open = df_hab$Area_Open,grassArea = df_hab$grassArea, shrubArea = df_hab$grassArea, treeArea = df_hab$treeArea, siteType = df_hab$Site_Type,month = month(df_hab$EndDate), SiteType = df_hab$SiteTypeRgp ,df_hab_sel)

```



Ordinal logistic regression

```{r}

df_hab_sc$Complexity<-ordered(as.factor(df_hab_sc$Complexity), levels = c("0","1", "2", "3", "4"))


m1<-polr(Complexity ~ auc_an_min_bio + peak_bio + peak_anth, data  =  df_hab_sc)
m2<-polr(Complexity ~ peak_bio + peak_anth, data  =  df_hab_sc)
m3<-polr(Complexity ~ peak_bio, data  =  df_hab_sc)
m4<-polr(Complexity ~ peak_anth, data  =  df_hab_sc)
m5<-polr(Complexity ~ auc_an_min_bio, data  =  df_hab_sc)

AIC(m1,m2,m3,m4,m5)

```
Cumulative link mixed models from the ordinal package
```{r}

df_hab_sc$site<-as.factor(df_hab_sc$site)
df_hab_sc$meanTemp_f<-as.factor(df_hab_sc$meanTemp)

glob_clm<-clmm(Complexity ~ n_peaks_anth + n_peaks_bio + auc_an_min_bio + peak_anth + peak_bio + peak_dawn_anth + peak_dawn_bio + (1|site), data = df_hab_sc)

summary(glob_clm)
```

```{r}
gc1<-update(glob_clm, ~ . -n_peaks_anth)
gc2<-update(gc1, ~ . -n_peaks_bio)
gc3<-update(gc2, ~ . -peak_anth)
gc4<-update(gc3, ~ . -peak_dawn_anth)
gc5<-update(gc4, ~ . -peak_dawn_bio)
gc6<-update(gc5, ~ . -peak_bio)

AIC(glob_clm, gc1, gc2, gc3, gc4, gc5, gc6)

# 
# summary(gc4)
# summary(gc5)
# summary(gc6)
```

```{r, eval = FALSE, echo = FALSE}

df_hab_sc$site<-as.factor(df_hab_sc$site)

glob_clm_temp<-clmm(Complexity ~ n_peaks_anth + n_peaks_bio + auc_an_min_bio + peak_anth + peak_bio + peak_dawn_anth + peak_dawn_bio + (1|tempRange), data = df_hab_sc)


summary(glob_clm_temp)

gct1<-update(glob_clm_temp, ~ . -n_peaks_anth)
gct2<-update(gct1, ~ . -peak_dawn_anth)
gct3<-update(gct2, ~ . -peak_bio)
gct4<-update(gct3, ~ . -n_peaks_bio)
gct5<-update(gct4, ~ . -peak_dawn_bio)
gct6<-update(gct5, ~ . -peak_anth)

AICtab(glob_clm_temp, gct1, gct2, gct3, gct4, gct5, gct6)

summary(gct2)
summary(gct6)
#pcor(df_hab_sc[,c(5:13)], method = "pearson")


```



```{r, eval = FALSE,echo=FALSE}
clm1<-clmm(Complexity~ auc_an_min_bio +(1|site) , data  =  df_hab_sc)
clm2<-clmm(Complexity ~ peak_anth + peak_bio + (1|site), data = df_hab_sc)
clm3<-clmm(Complexity ~ peak_bio + (1|site), data = df_hab_sc)
clm4<-clmm(Complexity ~ peak_anth + (1|site), data = df_hab_sc)
clm5<-clmm(Complexity ~ peak_anth + peak_bio + auc_an_min_bio + (1|site), data = df_hab_sc)


AICtab(glob_clm, clm1, clm2, clm3, clm4, clm5)
```



```{r, echo = FALSE}
# 
# glm1<-glm(siteDiversity ~ auc_an_min_bio+peak_bio+peak_anth, data  =  df_hab, family = poisson)
# glm2<-glm(siteDiversity ~ peak_bio+peak_anth, data  =  df_hab, family = poisson)
# glm3<-glm(siteDiversity ~ peak_bio, data  =  df_hab, family = poisson)
# glm4<-glm(siteDiversity ~ peak_anth, data  =  df_hab, family = poisson)
# glm5<-glm(siteDiversity ~ auc_an_min_bio, data  =  df_hab, family = poisson)
# 
# AIC(glm1,glm2,glm3,glm4,glm5)

```

```{r}

global<-glmer(siteDiversity ~ n_peaks_anth + n_peaks_bio + auc_an_min_bio + peak_anth + peak_bio + peak_dawn_anth + peak_dawn_bio + (1|site) , data = df_hab_sc, family = poisson)

glmm1<-glmer(siteDiversity ~ auc_an_min_bio+peak_bio+peak_anth + (1|site), data  =  df_hab_sc, family = poisson)
glmm2<-glmer(siteDiversity ~ peak_bio + peak_anth + (1|site), data  =  df_hab_sc, family = poisson)
glmm3<-glmer(siteDiversity ~ peak_bio + (1|site), data  =  df_hab_sc, family = poisson)
glmm4<-glmer(siteDiversity ~ peak_anth + (1|site), data  =  df_hab_sc, family = poisson)
glmm5<-glmer(siteDiversity ~ auc_an_min_bio + (1|site), data  =  df_hab_sc, family = poisson)

AIC(global, glmm1,glmm2,glmm3,glmm4,glmm5)



```


```{r}

library(bbmle)

#df_hab_sc<-scale(df_hab[,vars])

df_hab_sc$siteDiversity <-as.numeric(df_hab_sc$siteDiversity )

global<-glmer(siteDiversity  ~ n_peaks_anth + n_peaks_bio + auc_an_min_bio + peak_anth + peak_bio + peak_dawn_anth + peak_dawn_bio + (1|site), data = df_hab_sc, family = poisson)

glmm1<-glmer(siteDiversity ~ auc_an_min_bio+peak_bio+peak_anth + (1|site), data  =  df_hab_sc, family = poisson)
glmm2<-glmer(siteDiversity ~ peak_bio + peak_anth + (1|site), data  =  df_hab_sc, family = poisson)
glmm3<-glmer(siteDiversity ~ peak_bio + (1|site), data  =  df_hab_sc, family = poisson)
glmm4<-glmer(siteDiversity ~ peak_anth + (1|site), data  =  df_hab_sc, family = poisson)
glmm5<-glmer(siteDiversity ~ auc_an_min_bio + (1|site), data  =  df_hab_sc, family = poisson)

AICtab(global, glmm1,glmm2,glmm3,glmm4,glmm5)

```
```{r}

global_glm<-glm(treeArea + shrubArea + grassArea  ~ auc_bio + auc_anthro + SiteType + as.factor(month),  data = df_hab_sc, family = poisson)


global_glmer<-glmer(treeArea + shrubArea + grassArea  ~ auc_bio + auc_anthro + (1|SiteType) ,  data = df_hab_sc, family = poisson)

global_glmer_peak<-glmer(treeArea + shrubArea + grassArea  ~ peak_bio + peak_anth + (1|SiteTypeRgp),  data = df_hab, family = poisson)


AICtab(global_glmer, global_glmer_peak)

```

```{r,eval=FALSE, echo =  FALSE}

df_hab_sc$siteDiversity <-as.numeric(df_hab_sc$siteDiversity )

global_glm<-glm(siteDiversity  ~ auc_an_min_bio + peak_anth + peak_bio, data = df_hab_sc, family = poisson)

glm1<-glm(siteDiversity ~ auc_an_min_bio+peak_bio+peak_anth, data  =  df_hab_sc, family = poisson)
glm2<-glm(siteDiversity ~ peak_bio + peak_anth, data  =  df_hab_sc, family = poisson)
glm3<-glm(siteDiversity ~ peak_bio, data  =  df_hab_sc, family = poisson)
glm4<-glm(siteDiversity ~ peak_anth, data  =  df_hab_sc, family = poisson)
glm5<-glm(siteDiversity ~ auc_an_min_bio, data  =  df_hab_sc, family = poisson)


AICtab(global_glm, glm1,glm2,glm3,glm4,glm5)



```

```{r,eval=FALSE, echo =  FALSE}


global_glm<-glm(treeArea + shrubArea + grassArea  ~ auc_bio + auc_anthro + peak_dawn_anth + peak_dawn_bio, data = df_hab_sc, family = poisson)

glm1<-glm(treeArea + shrubArea + grassArea ~ auc_an_min_bio+peak_bio+peak_anth, data  =  df_hab_sc, family = poisson)
glm2<-glm(treeArea + shrubArea + grassArea ~ peak_bio + peak_anth, data  =  df_hab_sc, family = poisson)
glm3<-glm(treeArea + shrubArea + grassArea ~ peak_bio, data  =  df_hab_sc, family = poisson)
glm4<-glm(treeArea + shrubArea + grassArea ~ peak_anth, data  =  df_hab_sc, family = poisson)
glm5<-glm(treeArea + shrubArea + grassArea ~ auc_an_min_bio, data  =  df_hab_sc, family = poisson)


AICtab(global_glm, glm1,glm2,glm3,glm4,glm5)



```



Area Garden


Here we use the beta distribution as the response variable is a proportion - value between 0 and 1.  

```{r}

library(glmmTMB)
library(betareg)
library(glmmADMB)

df_hab$Area_Garden_Prop<-df_hab$Area_Garden/(pi*(100^2))
df_hab_sc$Area_Garden_Prop<-df_hab_sc$Area_Garden/(pi*(100^2))


df_hab$Area_Open_Prop<-df_hab$Area_Open/(pi*(100^2))
df_hab_sc$Area_Open_Prop<-df_hab_sc$Area_Open/(pi*(100^2))

df_hab$Area_Open_Garden_Prop<-df_hab$Area_Open_Prop+ df_hab$Area_Garden_Prop
df_hab_sc$Area_Open_Garden_Prop<-df_hab_sc$Area_Open_Prop+ df_hab_sc$Area_Garden_Prop

global_garden_admb<-glmmadmb(Area_Open_Garden_Prop ~   meanTemp +auc_an_min_bio + (1|site),  data = df_hab_sc, family="beta", link = "logit")

g1<-glmmadmb(Area_Open_Garden_Prop ~  auc_an_min_bio + (1|site),  data = df_hab_sc, family="beta")

gnull<-glmmadmb(Area_Open_Garden_Prop ~  meanTemp + (1|site),  data = df_hab_sc, family="beta")

AICtab(global_garden_admb, g1, gnull)


exp(cbind(OR = coef(g1), confint(g1)))


exp(cbind(OR = coef(global_garden_admb), confint(global_garden_admb)))


```

#global_garden_tmb<-glmmTMB(Area_Open_Garden_Prop ~   meanTemp + auc_an_min_bio  ,  data = df_hab_sc, family = list(family = "beta", link = "logit"))


```{r}
b1<-betareg(Area_Open_Garden_Prop ~ auc_an_min_bio +meanTemp, data = df_hab_sc, link = "logit")
b2<-betareg(Area_Open_Garden_Prop ~ auc_an_min_bio , data = df_hab_sc, link = "logit")
bnull<-betareg(Area_Open_Garden_Prop ~ meanTemp , data = df_hab_sc, link = "logit")


AICtab(b1, b2, bnull)

exp(cbind(OR = coef(b2), confint(b2)))



ran_gg<-as.data.frame(ranef(global_garden_admb))

ran_gg<-as.data.frame(ranef(g1))

ran_gg$site<-row.names(ran_gg)

ran_gg %>% 
  arrange(X.Intercept.)

df<-unique(merge(ran_gg, df_hab_sc[,c("site", "Area_Open_Garden_Prop")], by = "site"))



plot(df$X.Intercept., df$Area_Open_Garden_Prop)


# dotplot()
# 
# r.squaredGLMM(global_garden)

```

#Open and Garden together



```{r}

df_hab_sc$meanTemp_fac<-as.factor(df_hab_sc$meanTemp)

global_garden_rt<-glmmadmb(Area_Garden_Prop ~   auc_an_min_bio +(1|meanTemp_fac) ,  data = df_hab_sc, family="beta")


summary(global_garden_rt)

```

```{r}
glmm1<-glmer(siteDiversity ~ auc_an_min_bio+peak_bio+peak_anth + (1|site), data  =  df_hab_sc, family = poisson)
glmm2<-glmer(siteDiversity ~ peak_bio + peak_anth + (1|site)+ (1|meanTemp)+ (1|tempRange), data  =  df_hab_sc, family = poisson)
glmm3<-glmer(siteDiversity ~ peak_bio + (1|site)+ (1|meanTemp)+ (1|tempRange), data  =  df_hab_sc, family = poisson)
glmm4<-glmer(siteDiversity ~ peak_anth + (1|site)+ (1|meanTemp)+ (1|tempRange), data  =  df_hab_sc, family = poisson)
glmm5<-glmer(siteDiversity ~ auc_an_min_bio + (1|site)+ (1|meanTemp)+ (1|tempRange), data  =  df_hab_sc, family = poisson)

AIC(global, glmm1,glmm2,glmm3,glmm4,glmm5)

```


```{r}

df_hab_sc %>% 
  group_by(site) %>% 
  summarise(mean_auc_diff = mean(auc_an_min_bio), mean_area_green = mean(Area_Open_Garden_Prop)) %>% 
  ggplot(aes(x = mean_area_green, y = mean_auc_diff))+
  geom_point()+
  geom_smooth()



sum_df_hab_sc<-df_hab_sc %>% 
  group_by(site) %>% 
  summarise(mean_auc_diff = mean(auc_an_min_bio), mean_area_green = mean(Area_Open_Garden_Prop))


bmean<-betareg(sum_df_hab_sc$mean_area_green ~  sum_df_hab_sc$mean_auc_diff )


summary(bmean)

exp(cbind(OR = coef(bmean), confint(bmean)))



```


```{r}





```